### 2017-12-12 readme.rmd
added details on focal analyses

cteno_color = "#F8766D"
poro_color = "#00BA38"

@@

D %>%
ggplot( aes( Recoding, Dataset ) ) + 
	geom_tile( aes( fill = Result ) ) + 
	facet_wrap( ~ Model ) +
	ggtitle("Animal Root Support") +
	theme_classic()

@@

## Sensitivity and the support landscape

@@

We can think of this variation as a support landscape (Figure 4) - a multidimensional space that shows how support varies across different gene sampling, taxon sampling, and analysis methods. It is only ever possible to probe a finite region of this landscape, but its overall shape has radical implications for how we interpret a particular set of results. This means that when investigating sensitivity it is necessary to be explicit not only about the results we get in a given study, but how we are thinking about regions of the support landscape that are adjacent to those we examine.

The excerpts from Feuda *et al.* suggest that the authors implicitly envision the broader support landscape to be a monotonic surface (Figure 4a). They believe that the small part of the support landscape they probed (Figure 4, inset boxes) indicates a broader pattern that they interpret as strong support for Porifera-sister. This is because, although most of their analyses support Ctenophora-sister there are some analyses on the right edge of the sampled space that recover significant support for Porifera-sister.

![overview](figures/figure_4.png)
*Figure 4 - Hypothetical support landscapes. The inset box shows the region of the landscape explored by Figures 2-3. The horizontal axis represents variation in analysis methods, such as model of molecular evolution and recoding of data. The vertical axis represents variation in sampling of genes and species. (a) A Monotonic Support Landscape. (b) A Sponge-Island Support Landscape, which would indicate that sponges support is only obtained in a very small portion of the support landscape. (c) A Hilly Support Landscape.*

Other support landscapes would have very different implications for the results they obtained. A small fraction of analyses may have supported Porifera-sister because they happened to sample a small island of support (Figure 4b) in a broader landscape that overall has patterns of support consistent with Ctenophora-sister. Alternatively, the small island of support for Porifera-sister may indicate a transition from a plateau of support for Ctenophora sister to a hilly support landscape to the east. It is unlikely that we would conclusively resolve the placement of ctenophores and sponges if this broader hilly pattern were found to be the case.

>>>>

D$Maxdiff = round( D$Maxdiff, 2 )

cteno_color = "#F8766D"
poro_color = "#00BA38"
unresolved_color = "#6F94CD"

@@

gg_sig = 
	D %>%
		ggplot( aes( Recoding, Dataset ) ) + 
			geom_tile( aes( fill = Result ) ) + 
			facet_wrap( ~ Model ) +
			ggtitle("Animal Root Support") +
			theme_classic()

ggsave( "figures/figure_3.pdf", gg_sig, device=pdf, width=6.5, height=3.1 )

gg_sig

@@

## Sensitivity within the presented results

The primary explanation Feuda *et al.* 2017 provide for accepting some results and rejecting others is their posterior predictive (PP) analyses of model adequacy. They calculated PP for only the Whelan-Opisthokont and Chang matrices (Figures 2-3, top two rows). Based on these scores they rejected the results of the GTR-G analyses (Figures 2-3, left pane) and analyses that had not been recoded. This left six cells that are the focus of their primary conclusions (Figures 2-3, right panel, top two rows, right 3 cells).

```{r result_sub}

D_sub = 
	D %>%
	rename( Porifera_sister_posterior = Porifera_sister ) %>%
	filter(Model=="CAT-GTR+G") %>% 
	filter( (Dataset=="Whelan-Opis") | (Dataset=="Chang-Holo") ) %>% 
	filter( Recoding != "None" ) %>%
	select (-c(Ctenophora_sister, Burnin, effsize, rel_diff)) %>% 
	gather( var, val, Porifera_sister_posterior:PPA_MEAN ) %>% 
	mutate( var = factor( var, levels = c("Porifera_sister_posterior", "Cycles", "Maxdiff", "PPA_DIV", "PPA_MAX", "PPA_CONV", "PPA_VAR", "PPA_MEAN")))

gg_sub = 
	D_sub %>%
		ggplot( aes( Recoding, Dataset ) ) + 
			geom_tile( aes( fill = Result ) ) + 
			facet_wrap( ~ var, ncol = 2 ) +
			geom_text(aes(label = val)) +
			scale_fill_manual( values=c( poro_color,unresolved_color ) ) +
			ggtitle("Variation between analyses") +
			theme_classic()

ggsave( "figures/figure_4_raw.pdf", gg_sub, device=pdf, width=6.5, height=6 )

![figure 4](figures/figure_4.png)

*Figure 4 - Detailed examination of the analysis subset (indicated in upper right pane) that is the focus of the conclusions by Feuda et al. These are the six recoded CAT-GTR+G analyses with posterior predictive (PP) scores. The presented statistics for these six cells are posterior probability of Poifera-sister, Maxdiff (with lower scores indicating better convergence), and five PP scores (where lower absolute value indicates better model adequacy). Only one analysis provides significant support for Porifera-sister. It isn't the most adequate by any of the PP scores, and showed the poorest convergence according to Maxdiff. Recodings are: None- K - KGB-6, D - Dayhoff-6, and S- SR-6.*

A variety of statistics from Feuda *et al.* 2017 for these six cells are presented in Figure 4. The first thing to note is that only one of these six analyses provides significant support for Porifera-sister. This is for the SR-6 recoded Whelan-Opisthokont matrix, shown in green in Figure 4. Given the similarity of these six analyses, the fact that only one supports Porifera-sister calls into serious question the robustness of support for Porifera-sister.

The other statistics help put this single significant result into context, and they could be used to provide explicit criteria for favoring the one significant result over the five insignificant results. Feuda *et al.* present PP as the primary criterion we should consider. The one analysis with significant support for Porifera-sister does not have the best (lowest absolute value) PP score for any of the five PP statistics considered (Figure 4). By their own criterion of model adequacy, we should favor the unresolved analyses.

In fact, the only statistic that stands out for the significant analysis is the high Maxdiff score, which indicates it is the most poorly converged of these six focal analyses.

## Sensitivity and the broader support landscape

@@

We can think of this variation as a support landscape (Figure 5) - a multidimensional space that shows how support varies across different gene sampling, taxon sampling, and analysis methods. It is only ever possible to probe a finite region of this landscape, but its overall shape has radical implications for how we interpret a particular set of results. This means that when investigating sensitivity it is necessary to be explicit not only about the results we get in a given study, but how we are thinking about regions of the support landscape that are adjacent to those we examine.

The excerpts from Feuda *et al.* suggest that the authors implicitly envision the broader support landscape to be a monotonic surface (Figure 5a). They believe that the small part of the support landscape they probed (Figure 5, inset boxes) indicates a broader pattern that they interpret as strong support for Porifera-sister. This is because, although most of their analyses support Ctenophora-sister there are some analyses on the right edge of the sampled space that recover significant support for Porifera-sister.

![overview](figures/figure_5.png)
*Figure 5 - Hypothetical support landscapes. The inset box shows the region of the landscape explored by Figures 2-3. The horizontal axis represents variation in analysis methods, such as model of molecular evolution and recoding of data. The vertical axis represents variation in sampling of genes and species. (a) A Monotonic Support Landscape. (b) A Sponge-Island Support Landscape, which would indicate that sponges support is only obtained in a very small portion of the support landscape. (c) A Hilly Support Landscape.*

Other support landscapes would have very different implications for the results they obtained. A small fraction of analyses may have supported Porifera-sister because they happened to sample a small island of support (Figure 5b) in a broader landscape that overall has patterns of support consistent with Ctenophora-sister. Alternatively, the small island of support for Porifera-sister may indicate a transition from a plateau of support for Ctenophora sister to a hilly support landscape to the east. It is unlikely that we would conclusively resolve the placement of ctenophores and sponges if this broader hilly pattern were found to be the case.
